### 2009-09-09 latex/paper1/contents.tex
paper

Many models were proposed to explain the formation of scale-free networks. These
models are simple algorithms that can be proven, either formally or empirically,
to generate networks that are scale-free.

Glossary: preferential attachment

@@

In common: the were proven formally to generate scale-free directed networks.

@@

We propose an extension to BCR model... which was proposed to model the links
between web pages.
\cite{Bollobas2003}

@@

\section{Experimental Setup}
% TODO: use simple past tense

@@

@@

We already know that they
produce networks that, just like software networks, are scale-free. This single
property, though, isn't enough to prove the hypothesis, since there are many
known scale-free networks extracted from many distinct domains of knowledge.

@@

\subsection{Real Networks Data Set}

We have collected 131 networks to support the experiment. The networks are
described in detail in Appendix A.

Software networks. We have collected 65 software systems written in Java. Java
is a popular programming language, for which there is plenty of open source
software.  The corresponding software networks were extracted by the tool
Dependency Finder CITE, considering classes and interfaces as components.
Dependency Finder was used because it can be easily integrated with scripts.

Non-software networks. We have collected 66 networks from distinct domains, such
as biology, sociology, technology, and linguistics. These networks are freely
available on the Internet and have been used in previous researches.

%We collected 65 software systems written in the Java
%programming language and 66 networks from other domains, such as biology,
%sociology and linguistics. Both the software systems and the networks are freely
%available on the Internet.
%
%As of this writing, Java is a popular programming language, for which there is
%plenty of open source software CITE. Also, there are many dependency extractors
%for this language. We have extracted the dependency network for each software
%system using a tool called Dependency Finder CITE, for its easy integration with
%scripts. Dependency Finder looks for dependencies by means of static analysis of
%Java bytecode and outputs the dependency network in a XML file. Although this
%network is in a detailed level, showing dependencies between fine-grained
%implementation entities such as attributes and methods, we have abstracted it to
%a network of dependencies between classes. The network is directed and
%non-weighted. The size of the networks range from XXX to YYY.

% JARs?

%The networks from other domains were collected in websites of complex network
%research groups. The domains include sociology, linguistics, biology and
%technology. The size of the networks range from XXX to YYY vertices.

>>>>

Many models were proposed to explain the formation of scale-free networks. These
models are simple algorithms that can be proven, either formally or empirically,
to generate networks that are scale-free. In this section we present three
models that generate directed networks with built-in modular decomposition:
BCR+, CGW and LR. The first two models grow networks using a mechanism called
preferential attachment CITE Barabasi, by which nodes with many edges tend to
receive more edges in the process.

@@

@@

@@

\section{Characterization of Software Networks}

Our research hypothesis is that at least one of the presented models can
synthesize networks that resemble software networks. A central issue, thus, is
how to measure similarity between networks. In order to be useful, the metric
must be able to differentiate between software and non-software networks. In
this section we present such a metric, together with an experiment that
evaluates its usefulness by applying it to both software and non-software
networks. %Then we show a classification model

\subsection{A Metric for the Similarity Between Networks}

In a recent work, Milo et al. \cite{Milo2002} proposed to characterize networks
by analyzing their triad concentration. A triad is a network with three vertices
in which all vertices are connected. There are only 13 distinct triads, one for
each configuration of directed edges, as shown in Figure \ref{fig:triads}.

By counting how many times each triad appears in a network, one can build a
triad concentration profile (TCP), which is a vector with 13 numbers that
summarize the local structure of the network. Figure \ref{fig:profiles} shows
the TCP for three distinct networks.

Similarity between two networks can be measured by computing Pearson's
correlation coefficient between the corresponding TCPs, which yields a value
between -1 (most dissimilar) and 1 (most similar). This is the similarity metric
used in this research.

\subsection{Data Set}

To support the evaluation of the metric, we have collected 131 networks. The
networks are described in detail in Appendix A.

\textbf{Software networks}. We have collected 65 software systems written in
Java, with size ranging from 111 to 35,363 classes. Java was chosen for being a
popular programming language in which many open source systems were written. The
software networks, representing dependencies between classes, were extracted
with the tool Dependency Finder CITE.

\textbf{Non-software networks}. We have collected 66 networks from distinct
domains, such as biology, sociology, technology, and linguistics. These networks
are freely available on the Internet and have been used in previous researches.

% S-score

\subsection{Evaluation of the Similarity Metric}

In order to evaluate the similarity metric, we measured the similarity between
the networks in the data set. A suitable metric must fullfil two conditions: (i)
it must yield high similarity between software networks, and (ii) it must yield
lower similarity between software networks and networks from other domains.

We measured the similarity between each pair of software networks and then
calculated the average and the standard deviation. The average similarity was
0.97 and the standard deviation, 0.03. The high average similarity and the low
standard deviation show that software networks have common structural patterns
that are captured by the metric. The minimum similarity was 0.83.

TODO: Boxplot by network class (software, metabolic, neural, social etc.)

Then, for each non-software network, we measured the average similarity between
the network and all software networks. The majority of the networks (X\%) had an
average similarity lower than 0.83, which shows that they are ... Networks
extracted from the social network Facebook had negative similarities.

A few networks, though, showed high similarity to software networks. The network
polblogs, showing links between blogs on politics, had similarity 0.97. The
neural network of C Elegans also had a high similarity (0.88). Further
investigation is needed in order to discover why does it happen and whether
auxiliary metrics can differentiate these networks from software networks.

TODO: histogram of similarities? (maybe boxplot is enough)

\subsection{A Network Classification Model}

Although the similarity metric tells how close a network is from software
networks, it does not tell whether a network is close enough that it can be
considered a software network. Therefore, in this section we present a
classification model that splits networks into two classes: software-like
networks and non software-like networks.

We chose a simple classification model based on the similarity metric. We need
to define a similarity threshold, so networks with similarity below this
threshold are classified as non software-like, and networks with similarity
greater or equal are classified as software-like.

Classification models are subject to two types of errors: false negatives
(software networks that are classified as non software-like) and false positives
(non-software networks that are classified as software-like).

The choice of a threshold is arbitrary.
We want to maximize precision.

\section{Evaluation of Network Models}

In the previous section it was shown that many networks, although scale-free,
can be distinguished from software networks by a triad-based metric. In this
section we show empirically that the three network models previously presented
synthesize networks that are indistinguishable from software networks.

% TODO: use simple past tense

@@

A central issue is what

@@

% TODO: move to the section on the oracle
We already know that they
produce networks that, just like software networks, are scale-free. This single
property, though, isn't enough to prove the hypothesis, since there are many
known scale-free networks extracted from many distinct domains of knowledge.

@@
